---
title: "HW3-Solution"
author: "Yashashri Kulkarni"
date: "9/25/2021"
output: html_document
---
#### 1. Assume Z has a standard normal distribution. Determine the value for z that solves each of the following:
#### (a) P(−z<Z<z)=0.9

To determine z such that P (−z < Z ≤ z) = 0.9
If 90% of the area lies between −z and z, then 10% of the area must lie outside of this range. 
Since normal curves are symmetric, half of this amount– 5%– must lie before −z. 
Then the area under the curve before z must be: 0.05+0.9= 0.95   

```{r echo=TRUE}
z= qnorm(c(.05, .95))
z
```

#### (b) P(−1.1<Z <z)=0.8
80% of the area under the curve must lie between -1.1 and z.  
    => pnorm(z)- pnorm(-1.1)= 0.8  
    => pnorm(z)= 0.8+pnorm(-1.1)

```{r setup, echo=TRUE}
z= qnorm(0.8+pnorm(-1.1))
z
```
#### 2. The speed of a file transfer from a server on campus to a personal computer at a student’s home on a weekday evening is normally distributed with a mean of 60 kilobits per second and a standard deviation of 4 kilobits per second.
#### (a) What is the probability that the file will transfer at a speed of 70 kilobits per second or more?
```{r echo= TRUE}
mu <- 60 ; sigma <- 4
#P(z> 70)
1- (pnorm(70, mu, sigma)- pnorm(0, mu, sigma))
```
#### (b) What is the probability that the file will transfer at a speed of less than 58 kilobits per second?
``` {r echo= TRUE}
#P(0< z< 58)
(pnorm(58, mu, sigma)- pnorm(0, mu, sigma))
```
#### (c) If the file is 1 megabyte, what is the average time it will take to transfer the file? (Assume eight bits per byte.)
```{r echo= TRUE}
size= 1000*8 #in kilobits
t= size/mu
t/60 #in minutes  
```

#### 3. Hits to a high-volume Web site are assumed to follow a Poisson distribution with a mean of 10,000 per day. Approximate each of the following:
#### (a) The probability of more than 20,000 hits in a day
```{r echo= TRUE}
ppois(20000, lambda= 10000,lower.tail = FALSE)
```
#### (b) The probability of less than 9900 hits in a day
```{r echo= TRUE}
ppois(9900, lambda= 10000,lower.tail = TRUE)
```

#### (c) The value such that the probability that the number of hits in a day exceed the value is 0.01
```{r echo= TRUE}
qpois(.01, lambda = 10000, lower.tail = FALSE)
```
#### 4. Calls to the help line of a large computer distributor follow a Poisson distribution with a mean of 20 calls per minute.
#### (a) What is the mean time until the one-hundredth call? (3 points)
```{r echo= TRUE}
# The Erlang random variable describes the time interval between any event and the kth following event.
# mean of Erlang RV= k/lambda
# Here r= 100, lambda= 20
100/20
```
#### (b) What is the mean time between call numbers 50 and 80? (3 points)
```{r echo= TRUE}
# r= 30, lambda= 20
30/20
```
#### (c) What is the probability that three or more calls occur within 15 seconds? (5 points)
P(X>= 3)= 1-[P(0)+P(1)+P(2)]
```{r echo= TRUE}
# Creating a function to calculate the Erlang pdf using the given parameters altogether
# Here lambda,l, i.e.the mean, is taken as 5. 
# (20 calls per minute== 5 calls per 15 seconds)

1-ppois(2, 5)
```
#### 5. Suppose the random variables X, Y, and Z have the following joint probability distribution.
Determine the following: 
```{r echo= FALSE}
# creating the joint probability matrix
x= c(1,1,1,1,2,2,2,2)
y= c(1,1,2,2,1,1,2,2)
z= c(1,2)
f_XYZ= c(.05,.1, .15, .2,.2, .15, .1, .05)
p= cbind(x,y,z,f_XYZ)
p.df= as.data.frame(p)
p.df
```
#### (a) P(X = 2)
```{r echo= TRUE}
#For X=2
sum(f_XYZ[x==2])
```
#### (b) P(X=1,Y =2) 
```{r echo= TRUE}
sum(subset.data.frame(p.df, p.df$x==1 & p.df$y== 2)$f_XYZ)
```
#### (c) P(Z < 1.5)
```{r echo= TRUE}
sum(subset.data.frame(p.df, p.df$z< 1.5)$f_XYZ)
```
#### (d) P(X=1orZ=2)
```{r echo= TRUE}
sum(subset.data.frame(p.df, p.df$x==1 | p.df$z== 2)$f_XYZ)
```
#### (e) E(X)
```{r echo= TRUE}
mu_X= sum(x*f_XYZ)
mu_X
```
#### (f) P(X = 1|Y = 1)
```{r echo= TRUE}
sum(subset.data.frame(p.df, p.df$x==1 & p.df$y== 1)$f_XYZ)/ sum(subset.data.frame(p.df, p.df$y== 1)$f_XYZ)
```
#### (g) P(X =1,Y =1|Z =2)
```{r echo= TRUE}
sum(subset.data.frame(p.df, p.df$x==1 & p.df$y== 1 & p.df$z== 2)$f_XYZ)/ sum(subset.data.frame(p.df, p.df$z== 2)$f_XYZ)
```
#### (h) P(X =1|Y =1,Z =2)
```{r echo= TRUE}
sum(subset.data.frame(p.df, p.df$x==1 & p.df$y== 1 & p.df$z== 2)$f_XYZ)/ sum(subset.data.frame(p.df, p.df$y== 1 & p.df$z== 2)$f_XYZ)
```
#### (i) Conditional probability distribution of X given that Y = 1 and Z = 2.
```{r echo= TRUE}
# When X=1, Y=1, Z= 2
sum(subset.data.frame(p.df, p.df$x==1 & p.df$y== 1 & p.df$z== 2)$f_XYZ)/sum(subset.data.frame(p.df, p.df$y== 1 & p.df$z== 2)$f_XYZ)
# When X=2, Y=1, Z= 2
sum(subset.data.frame(p.df, p.df$x==2 & p.df$y== 1 & p.df$z== 2)$f_XYZ)/sum(subset.data.frame(p.df, p.df$y== 1 & p.df$z== 2)$f_XYZ)
```
#### (j) Cov(x,y) 
```{r echo= TRUE}
mu_Y= sum(y*f_XYZ)
mu_Y
cov_XY= sum((x-mu_X)*(y-mu_Y)*f_XYZ)
cov_XY
```
#### (k) Cov(x,z) 
```{r echo= TRUE}
mu_Z= sum(z*f_XYZ)
mu_Z
cov_XZ= sum((x-mu_X)*(z-mu_Z)*f_XYZ)
cov_XZ
```
#### (l) Cov(y,z)
```{r echo= TRUE}
cov_YZ= sum((y-mu_Y)*(z-mu_Z)*f_XYZ)
cov_YZ
```
```{r echo= TRUE}
library(pracma)
```
#### 6. Consider the joint probability density function fXY (x,y) = c(x+y) over the range 0 < x < 3 and x<y<x+2. Determine
#### (a) c
```{r echo= TRUE}
fun= function(x,y){x+y}
ymin= function(x){x}
ymax= function(x){x+2}
i= integral2(fun,xmin= 0,xmax= 3,ymin,ymax)$Q
i
```
c= 1/i= 1/24  

#### (b) P(X<1,Y <2)
From above steps, f(x,y)= (x+y)/24
```{r echo= TRUE}
fun= function(x,y){(x+y)/24}
ymin= function(x){x}
integral2(fun,xmin= 0, xmax= 1,ymin, ymax=2)$Q
```
#### (c) P(1<X<2)
```{r echo= TRUE}
ymin= function(x){x}
ymax= function(x){x+2}

i= integral2(fun,xmin= 1,xmax= 2,ymin, ymax)$Q
i
```
#### (d) P(Y > 1)
```{r echo= TRUE}
ymin= function(x){x}
ymax= function(x){x+2}

a= integral2(fun, xmin= 0,xmax= 1, ymin= 1,ymax)$Q
b= integral2(fun, xmin= 1, xmax= 3,ymin,ymax)$Q
a+b
```
#### (e) P(X<2,Y <2) (f) E(X)
```{r echo= TRUE}
ymin= function(x){x}
i= integral2(fun, 0,2,ymin,2)$Q
print(i)
```
#### (f) E(X)
```{r echo= TRUE}
fun= function(x, y){x*(x+y)/24}
ymin= function(x){x}
ymax= function(x){x+2}
E_X= integral2(fun, 0,3,ymin, ymax)
E_X$Q
```

#### (g) Marginal probability distribution of X  
$\int_{x}^{x+2}\left(\frac{x+y}{24}\right) \;dy$  
= $\left.\frac{1}{24}\left(xy+\frac{y^2}{2}\right) \right|_{x}^{x+2}\;dx$   
= $\frac{1}{12} (2x+1)$

```{r echo= TRUE}
(2*(0:2)+1)/12  
```
#### (h) Conditional probability distribution of Y given that X = 1  
= $\frac {f(X=1, Y=y)}{f(X=1)}$  
= $\frac {y+1}{6}$  
```{r echo= TRUE}
(c(1:3)+1)/12 
```

#### (i) E(Y|X =1)
```{r echo= TRUE}
fun= function(y){y*(y+1)/6}
integral(fun, 1,3)
```
#### (j) P(Y >2|X =1)
```{r echo= TRUE}
fun= function(y){(y+1)/6}
integral(fun, 2,3)
```
#### (k) Conditional probability distribution of X given that Y = 2  
Conditional Prob= Joint Dist/Marginal Dist  
Joint.dist= $\int_{0}^{y}\left(\frac{1}{24} (x+y)\right) \;dx$ = $\frac {y^2}{16} $  
Marginal.dist= $\frac {(x+2)}{24}$  

Cond.Prob= $\frac {(x+2)/24}{2^2/16}= \frac {x+2}{6}$

#### (l) Cov(x,y) 
cov(x,y)= E(XY)- E(X)E(Y)
```{r echo= TRUE}
fun= function(x, y){y*(x+y)/24}
ymin= function(x){x}
ymax= function(x){x+2}
E_Y= integral2(fun, 0,3,ymin, ymax)

fun= function(x,y){x*y*(x+y)/24}
ymin= function(x){x}
ymax= function(x){x+2}
E_XY= integral2(fun, 0,3,ymin, ymax)

Cov_XY= E_XY$Q-(E_X$Q*E_Y$Q)
Cov_XY
```
#### (m) ρx,y

```{r echo= TRUE}
E_X$Q
fun= function(x, y){x^2*(x+y)/24}
ymin= function(x){x}
ymax= function(x){x+2}
E_X2= integral2(fun, 0,3,ymin, ymax)$Q

fun= function(x, y){y^2*(x+y)/24}
ymin= function(x){x}
ymax= function(x){x+2}
E_Y2= integral2(fun, 0,3,ymin, ymax)$Q

VarX= E_X2- (E_X$Q)^2
VarY= E_Y2- (E_Y$Q)^2

Cov_XY/(sqrt(VarX)* sqrt(VarY))
```