---
title: "Midterm- Solution"
author: "Yashashri Kulkarni"
date: "10/10/2021"
output: html_document
---
I have neither given nor recieved aid on this exam, nor have I concealed any violation of the Honor Code.

(a) (FALSE) 
(b) (FALSE) 
(c) (FALSE) 
(d) (TRUE) 
(e) (FALSE)
(f) (TRUE) 

Problem 2 (9 points total, 3 points each part)

```{r echo= TRUE}
## This is a binomial Dist with success p= .65 and failure, q= .35
#(a) hits the bull’s eye for the 10th time on the 15th try?
dbinom(9,14,.65)* dbinom(1,1,.65)

#(b) hits the bull’s eye 10 times in 15 tries?
dbinom(10 ,15, .65)
#(c) hits the first bull’s eye on the third try?
p=.65
q=.35
(p^1)*(q^2)
```

Problem 3 (9 points total, 3 points each part)
```{r echo= TRUE}
# A)Probability of person being a criminal, P(C)
#Using total probability law
(.3*.25)+ (.7*.1)
```
B) What is the probability that the person is from group 1 if he is a criminal?  
$P(A | C)= \frac {P(A \cap C)}{P(C)}$
```{r echo= TRUE}
(.3* .25)/.145
```

C) What is the probability that the person is from group 2 if she is a criminal?  
$P(B | C)= \frac {P(B \cap C)}{P(C)}$
```{r echo= TRUE}
(.7*.1)/ .145
```

Problem 4 (16 points total, 2 points each part)

1.CDF: discreet var  
2.Not a CDF  
3.Not a CDF  
4.Not a CDF  
5.CDF, Continous  
6.Not a CDF  
7.CDF, Continous  
8.CDF, Continous  


Problem 5 (18 points total, 3 points each part)
a) c
```{r echo= TRUE}
library(pracma)
x= c(0:2)
y= c(0:2)

fun= function (x, y){
  x^2+ (x*y)}
c= 1/integral2(fun,xmin=0, xmax=2, ymin= 0, ymax=2)$Q
c
```
b) Joint cdf $F_{X,Y}(x,y)$
\begin{aligned}
F_{X,Y}(x,y) &= \int_{-\infty}^{y}\int_{-\infty}^{x}{f_{X,Y}(x,y)dxdy}\\
&= \frac{3}{28}\int_{0}^{y}\int_{0}^{x}{(x^{2}+ xy)}dxdy\\
&= \frac{3}{28}\int_{0}^{y}{\left(\frac{x^{3}}{3}+ \frac{x^{2}y}{2}\right)}dy\\
&= \frac{1}{28*4}{\left(\frac{4x^{3}}{y}+ \frac{3x^{2}y^{2}}{2}\right)}\\
&= \frac{1}{112}{\left(\frac{4x^{3}}{y}+ \frac{3x^{2}y^{2}}{2}\right)}\\
\end{aligned}

```{r echo= TRUE}
f_XY = function(x,y){3/28*(x**2 + x*y)}
F_XY = integrate(function(x){
  sapply(x, function(x) {
    integrate(function(y) f_XY(x,y), 0, 2)$value
  })
}, 0, 2)$value
F_XY
```
c) Find the marginal cumulative distribution functions FX(x) and FY(y) and the marginal pdf fX(x) and fY(y) 

```{r echo= TRUE}
# Marginal Probability Distribution function of x
f_X = function(X){
  sapply(x, function(x){
    integrate(function(y) f_XY(x,y), 0, 2)$value
  })
}
f_X

# Marginal Probability Distribution function of y
f_y = function(X){
  sapply(x, function(x){
    integrate(function(y) f_XY(x,y), 0, 2)$value
  })
}
f_y

# CDF of x
F_X = integrate(function(x){
  sapply(x, function(x) {
    integrate(function(y) f_XY(x,y),0,2)$value
  })
},0,2)$value
F_X

# CDF of y
F_Y = integrate(function(x){
  sapply(x, function(x) {
    integrate(function(y) f_XY(x,y),0,2)$value
  })
},0,2)$value

F_Y
```
d) Find E(X) and Var(X)
```{r echo= TRUE}
# Mean and Variance of x
mu_x = integrate(function(x){
  sapply(x, function(x){
    integrate(function(y) x*f_XY(x,y), 0, 2)$value
  })
}, 0, 2)$value
mu_x

var_X = integrate(function(x){
  sapply(x, function(x){
    integrate(function(y) (x-mu_x)**2*f_XY(x,y), 0, 2)$value
  })
}, 0, 2)$value
var_X
```
e) Find E(Y ) and Var(Y)
```{r echo= TRUE}
# Mean and variance of y
mu_y = integrate(function(x){
  sapply(x, function(x){
    integrate(function(y) y*f_XY(x,y), 0, 2)$value
  })
}, 0, 2)$value
mu_y

var_y = integrate(function(x){
  sapply(x, function(x){
    integrate(function(y) (y-mu_y)**2*f_XY(x,y), 0, 2)$value
  })
}, 0, 2)$value
var_y
```

f) Find the covariance and correlation of X and Y 
```{r echo= TRUE}
# Covariance of X,Y
cov_XY = integrate(function(x){
  sapply(x,function(x) {
    integrate(function(y) (x-mu_x)*(y-mu_y)*f_XY(x,y),  0, 2)$value
  })
}, 0, 2)$value 
cov_XY

# correlation of X,Y
sigma_x = sqrt(integrate(function(x){
  sapply(x, function(x) {
    integrate(function(y) (x-mu_x)**2*f_XY(x,y), 0 ,2)$value
  })
}, 0, 2)$value)
sigma_x

sigma_y = sqrt(integrate(function(x){
  sapply(x, function(x) {
    integrate(function(y) (x-mu_y)**2*f_XY(x,y), 0 ,2)$value
  })
}, 0, 2)$value)
sigma_y

rho_XY = cov_XY/(sigma_x + sigma_y)
rho_XY
```
Problem 6. (a)
```{r echo= TRUE}
#C is type of computer randomly chosen
c= c(0,0,0,0,1,1,1,1)
#R is number of times rebooted
r= c(1,2,3,8,1,2,3,8)
f_cr= c(1/6, 2/6,0,1/6,0,1/6,1/6,0)
f_cr[c==1]

#finding the marginal dist of c
C =unique(c)
f_c= rep(0,length(C))
for (i in 1: length(C)){
  f_c[i]= sum(f_cr[c=C[i]])
}
f_c

R =unique(r)
f_r= rep(0,length(R))
for (i in 1: length(R)){
  f_r[i]= sum(f_cr[r=R[i]])
}
f_r
```
(b)E(C) and E(R) 
```{r echo= TRUE}
(mu_c= sum(c*f_cr))
(mu_r= sum(r*f_cr))

(mu_cr= sum(c*r*f_cr))
```

(c) Covariance of C and R
```{r echo= TRUE}
#Cov_CR
#Cov_CR= E(C)* E(R)- E(CR)
(Cov_CR= mu_c*mu_r-mu_cr)

#As the cov is positive, they are directly proportional and hence they are not independent.
```
(d) Joint Prob of M and W  
6,d,a) Create the table of joint prob and marginal pmfs.
```{r echo= FALSE}
m= c(1,2,8,1,2,8)
w= c(2,3,2,3,2,3)
f_MW= c(1/8, 2/8,1/8,1/8,2/8,1/8)
f_MW_df= data.frame(m,w,f_MW, stringsAsFactors = FALSE)
f_MW_df

f_w=c(c(sum(f_MW_df$f_MW[f_MW_df$w==c(2)])), c(sum(f_MW_df$f_MW[f_MW_df$w==c(3)])),sum(f_MW))
f_w

f_m= c(c(sum(f_MW_df$f_MW[f_MW_df$m==c(1)])),c(sum(f_MW_df$f_MW[f_MW_df$m==c(2)])),c(sum(f_MW_df$f_MW[f_MW_df$m==c(8)])),c(sum(f_MW_df$f_MW[f_MW_df$m==c(1)])),c(sum(f_MW_df$f_MW[f_MW_df$m==c(2)])),c(sum(f_MW_df$f_MW[f_MW_df$m==c(8)])),sum(f_MW))
f_m
t1= rbind(f_MW_df, f_w)
table= cbind(t1, f_m)
table
```
6,d,b) Calculate P(W>M)
```{r echo= TRUE}
1/8+1/8+2/8
```
6,d,c) Correlation
```{r echo= TRUE}
mu_W= 2*.5+ 3*.5
mu_M= 1*2/8 + 2*4/8+ 8*2/8

mu_WM= 2/8+3/8+1+12/8+16/8+24/8
mu_WM

Cov_WM= mu_W*mu_M- mu_WM
Cov_WM
```
Since Covariance is 0, correlation would be zero too.

Problem 8
```{r echo= TRUE}
y= c(293,230,172,91,113,125)
x1= c(1.6, 15.5, 22, 43, 33, 40)
x2= c(851,816,1058,1201,1357,1115)

b2= (sum(y*x1)*sum(x2^x1)- sum(x1^2)*sum(x2*y))/((sum(x1*x2)*sum(x2*x1))-(sum(x1^2)*sum(x2^2)))
b2

b1= (sum(x2^2)*sum(x1^y)- sum(x1*x2)*sum(x2*y))/((sum(x1^2)*sum(x2^2))-(sum(x1*x2)*sum(x2*x1)))
b1
```

